---
title: "Cut-points analysis"
author: "Lucas Veras"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: false
    toc_float: 
      collapsed: false
    theme: flatly
    code_folding: "show"
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load_packs}
library(here)
library(tidyverse)
library(knitr)
library(broman)
```


# Sample

```{r load_data}
source(here("code", "scripts", "01_clean_data.R"))
```

```{r sample_message}
if (sample == 1) {
  message <- "participants who have been evaluated on at least 2 of the 3 time-points (1, 6 and 12 months after surgery)"
} else {
  if (sample == 2) {
    message <- "participants who have been evaluated on the 3 time-points (1, 6 and 12 months after surgery), with training sessions attendance <= 50%"
  } else {
    if (sample == 3) {
      message <- "participants who have been evaluated on the 3 time-points (1, 6 and 12 months after surgery), no matter training sessions attendance"
    }
  }
}

n_eval <- hip %>% 
  filter(!is.na(AC)) %>% 
  group_by(eval) %>% 
  summarise(n = n_distinct(ID))

n <- hip %>% 
  filter(!is.na(AC)) %>% 
  summarise(n = n_distinct(ID))
```

The selected sample was all `r message`. The number of participants for the 1, 6 and 12 months after surgery evaluations was `r n_eval[1, 2]`, `r n_eval[2, 2]` and `r n_eval[3, 2]`, respectively.

Descriptive statistics for cardiorespiratory and accelerometry variables, separated by speed is show below, where `speed 0` is sitting rest, `speed 1` is standing rest, and the remaining `speed` values is the walking speed in km^.^h^-1^.

```{r explore}
source(here("code", "scripts", "02_explore.R"))
```

```{r cardio_acc_desc}
kable(
  cardio_acc_descriptives,
  col.names = c(
    "speed", "n",
    "kcal (mean)", "kcal (SD)",
    "VO~2~ (mean)", "VO~2~ (SD)",
    "MET (mean)", "MET (SD)",
    "AC (mean)", "AC (SD)",
    "ENMO (mean)", "ENMO (SD)",
    "MAD (mean)", "MAD (SD)"
  )
  )
```

AC, activity counts; ENMO, euclidean norm minus one; kcal, kilocalories; MAD, mean amplitude deviation; MET, metabolic equivalent; VO~2~, oxygen consumption.

# ROC curves

Cut-points for 1, 6 or 12 months after surgery that identify sedentary activity (SA) and physical activity intensities (PAI) created from AC, ENMO and MAD were obtained applying receiver operating characteristic curves (ROC). The indices used to summarize the cut-points were sensitivity (proportion of true positives that are correctly identified), specificity (proportion of false positive that are correctly identified) and the area under the curve.

```{r ROC}
source(here("code", "scripts", "03_create_ROC_curves.R"))
```

## 1 month after surgery

```{r ROC_1m}
ROC_summary_1m <- data.frame(
  acc_metric = c(rep("AC", 3), rep("ENMO", 3), rep("MAD", 3)),
  intensity = rep(c("Sedentary", "Moderate", "Vigorous"), 3),
  threshold = c(
    round(cp_AC_ROC_SED_1m[[1]], 0), round(cp_AC_ROC_MOD_1m[[1]], 0), round(cp_AC_ROC_VIG_1m[[1]], 0),
    round(cp_ENMO_ROC_SED_1m[[1]], 0), round(cp_ENMO_ROC_MOD_1m[[1]], 0), round(cp_ENMO_ROC_VIG_1m[[1]], 0),
    round(cp_MAD_ROC_SED_1m[[1]], 0), round(cp_MAD_ROC_MOD_1m[[1]], 0), round(cp_MAD_ROC_VIG_1m[[1]], 0)
  ),
  sensitivity = c(
    myround(cp_AC_ROC_SED_1m[[3]], 2), myround(cp_AC_ROC_MOD_1m[[3]], 2), myround(cp_AC_ROC_VIG_1m[[3]], 2),
    myround(cp_ENMO_ROC_SED_1m[[3]], 2), myround(cp_ENMO_ROC_MOD_1m[[3]], 2), myround(cp_ENMO_ROC_VIG_1m[[3]], 2),
    myround(cp_MAD_ROC_SED_1m[[3]], 2), myround(cp_MAD_ROC_MOD_1m[[3]], 2), myround(cp_MAD_ROC_VIG_1m[[3]], 2)
  ),
  specificity = c(
    myround(cp_AC_ROC_SED_1m[[2]], 2), myround(cp_AC_ROC_MOD_1m[[2]], 2), myround(cp_AC_ROC_VIG_1m[[2]], 2),
    myround(cp_ENMO_ROC_SED_1m[[2]], 2), myround(cp_ENMO_ROC_MOD_1m[[2]], 2), myround(cp_ENMO_ROC_VIG_1m[[2]], 2),
    myround(cp_MAD_ROC_SED_1m[[2]], 2), myround(cp_MAD_ROC_MOD_1m[[2]], 2), myround(cp_MAD_ROC_VIG_1m[[2]], 2)
  ),
  AUC = c(
    myround(auc(AC_ROC_SED_1m)[[1]], 2), myround(auc(AC_ROC_MOD_1m)[[1]], 2), myround(auc(AC_ROC_VIG_1m)[[1]], 2),
    myround(auc(ENMO_ROC_SED_1m)[[1]], 2), myround(auc(ENMO_ROC_MOD_1m)[[1]], 2), myround(auc(ENMO_ROC_VIG_1m)[[1]], 2),
    myround(auc(MAD_ROC_SED_1m)[[1]], 2), myround(auc(MAD_ROC_MOD_1m)[[1]], 2), myround(auc(MAD_ROC_VIG_1m)[[1]], 2)
  )
)
kable(
  ROC_summary_1m, 
  align = c("l", "l", "r", "r", "r", "r"), 
  col.names = c(
    "Accelerometer metric", "Intensity category", "Threshold", "Sensitivity", "Specificity", "AUC"
    )
  )
```

## 6 months after surgery

```{r ROC_6m}
ROC_summary_6m <- data.frame(
  acc_metric = c(rep("AC", 3), rep("ENMO", 3), rep("MAD", 3)),
  intensity = rep(c("Sedentary", "Moderate", "Vigorous"), 3),
  threshold = c(
    round(cp_AC_ROC_SED_6m[[1]], 0), round(cp_AC_ROC_MOD_6m[[1]], 0), round(cp_AC_ROC_VIG_6m[[1]], 0),
    round(cp_ENMO_ROC_SED_6m[[1]], 0), round(cp_ENMO_ROC_MOD_6m[[1]], 0), round(cp_ENMO_ROC_VIG_6m[[1]], 0),
    round(cp_MAD_ROC_SED_6m[[1]], 0), round(cp_MAD_ROC_MOD_6m[[1]], 0), round(cp_MAD_ROC_VIG_6m[[1]], 0)
  ),
  sensitivity = c(
    myround(cp_AC_ROC_SED_6m[[3]], 2), myround(cp_AC_ROC_MOD_6m[[3]], 2), myround(cp_AC_ROC_VIG_6m[[3]], 2),
    myround(cp_ENMO_ROC_SED_6m[[3]], 2), myround(cp_ENMO_ROC_MOD_6m[[3]], 2), myround(cp_ENMO_ROC_VIG_6m[[3]], 2),
    myround(cp_MAD_ROC_SED_6m[[3]], 2), myround(cp_MAD_ROC_MOD_6m[[3]], 2), myround(cp_MAD_ROC_VIG_6m[[3]], 2)
  ),
  specificity = c(
    myround(cp_AC_ROC_SED_6m[[2]], 2), myround(cp_AC_ROC_MOD_6m[[2]], 2), myround(cp_AC_ROC_VIG_6m[[2]], 2),
    myround(cp_ENMO_ROC_SED_6m[[2]], 2), myround(cp_ENMO_ROC_MOD_6m[[2]], 2), myround(cp_ENMO_ROC_VIG_6m[[2]], 2),
    myround(cp_MAD_ROC_SED_6m[[2]], 2), myround(cp_MAD_ROC_MOD_6m[[2]], 2), myround(cp_MAD_ROC_VIG_6m[[2]], 2)
  ),
  AUC = c(
    myround(auc(AC_ROC_SED_6m)[[1]], 2), myround(auc(AC_ROC_MOD_6m)[[1]], 2), myround(auc(AC_ROC_VIG_6m)[[1]], 2),
    myround(auc(ENMO_ROC_SED_6m)[[1]], 2), myround(auc(ENMO_ROC_MOD_6m)[[1]], 2), myround(auc(ENMO_ROC_VIG_6m)[[1]], 2),
    myround(auc(MAD_ROC_SED_6m)[[1]], 2), myround(auc(MAD_ROC_MOD_6m)[[1]], 2), myround(auc(MAD_ROC_VIG_6m)[[1]], 2)
  )
)
kable(
  ROC_summary_6m, 
  align = c("l", "l", "r", "r", "r", "r"), 
  col.names = c(
    "Accelerometer metric", "Intensity category", "Threshold", "Sensitivity", "Specificity", "AUC"
    )
  )
```

## 12 months after surgery

```{r ROC_12m}
ROC_summary_12m <- data.frame(
  acc_metric = c(rep("AC", 3), rep("ENMO", 3), rep("MAD", 3)),
  intensity = rep(c("Sedentary", "Moderate", "Vigorous"), 3),
  threshold = c(
    round(cp_AC_ROC_SED_12m[[1]], 0), round(cp_AC_ROC_MOD_12m[[1]], 0), round(cp_AC_ROC_VIG_12m[[1]], 0),
    round(cp_ENMO_ROC_SED_12m[[1]], 0), round(cp_ENMO_ROC_MOD_12m[[1]], 0), round(cp_ENMO_ROC_VIG_12m[[1]], 0),
    round(cp_MAD_ROC_SED_12m[[1]], 0), round(cp_MAD_ROC_MOD_12m[[1]], 0), round(cp_MAD_ROC_VIG_12m[[1]], 0)
  ),
  sensitivity = c(
    myround(cp_AC_ROC_SED_12m[[3]], 2), myround(cp_AC_ROC_MOD_12m[[3]], 2), myround(cp_AC_ROC_VIG_12m[[3]], 2),
    myround(cp_ENMO_ROC_SED_12m[[3]], 2), myround(cp_ENMO_ROC_MOD_12m[[3]], 2), myround(cp_ENMO_ROC_VIG_12m[[3]], 2),
    myround(cp_MAD_ROC_SED_12m[[3]], 2), myround(cp_MAD_ROC_MOD_12m[[3]], 2), myround(cp_MAD_ROC_VIG_12m[[3]], 2)
  ),
  specificity = c(
    myround(cp_AC_ROC_SED_12m[[2]], 2), myround(cp_AC_ROC_MOD_12m[[2]], 2), myround(cp_AC_ROC_VIG_12m[[2]], 2),
    myround(cp_ENMO_ROC_SED_12m[[2]], 2), myround(cp_ENMO_ROC_MOD_12m[[2]], 2), myround(cp_ENMO_ROC_VIG_12m[[2]], 2),
    myround(cp_MAD_ROC_SED_12m[[2]], 2), myround(cp_MAD_ROC_MOD_12m[[2]], 2), myround(cp_MAD_ROC_VIG_12m[[2]], 2)
  ),
  AUC = c(
    myround(auc(AC_ROC_SED_12m)[[1]], 2), myround(auc(AC_ROC_MOD_12m)[[1]], 2), myround(auc(AC_ROC_VIG_12m)[[1]], 2),
    myround(auc(ENMO_ROC_SED_12m)[[1]], 2), myround(auc(ENMO_ROC_MOD_12m)[[1]], 2), myround(auc(ENMO_ROC_VIG_12m)[[1]], 2),
    myround(auc(MAD_ROC_SED_12m)[[1]], 2), myround(auc(MAD_ROC_MOD_12m)[[1]], 2), myround(auc(MAD_ROC_VIG_12m)[[1]], 2)
  )
)
kable(
  ROC_summary_12m, 
  align = c("l", "l", "r", "r", "r", "r"), 
  col.names = c(
    "Accelerometer metric", "Intensity category", "Threshold", "Sensitivity", "Specificity", "AUC"
    )
  )
```

# Validity analysis

The validity of cut-points developed were posteriorly analyzed through leave-one-out cross-validation (LOOCV) method. Hence, each participant’s data was separated into a testing dataset (one at a time) with the remaining data being in the training dataset. New ROC curves, with the same structure as determined for the entire sample, were developed using the training dataset to identify SA or PAI for the only participant in the testing dataset. This process was repeated for all participants (`r n` times).

```{r validity}
source(here("code", "scripts", "04_analyse_ROC_validity.R"))
```

## Kappa statistic

### 1 month after surgery

```{r kappa_1m}
kappa_1m <- data.frame(
  acc_metric = c(rep("AC", 4), rep("MAD", 4), rep("ENMO", 4)),
  intens_cat = rep(c("Sedentary", "Moderate", "Vigorous", "Global"), 3),
  kappa = c(
    myround(kappa_AC_SED_1m[[5]], 2), myround(kappa_AC_MOD_1m[[5]], 2), myround(kappa_AC_VIG_1m[[5]], 2), myround(kappa_AC_glob_1m[[5]], 2),
    myround(kappa_ENMO_SED_1m[[5]], 2), myround(kappa_ENMO_MOD_1m[[5]], 2), myround(kappa_ENMO_VIG_1m[[5]], 2), myround(kappa_ENMO_glob_1m[[5]], 2),
    myround(kappa_MAD_SED_1m[[5]], 2), myround(kappa_MAD_MOD_1m[[5]], 2), myround(kappa_MAD_VIG_1m[[5]], 2), myround(kappa_MAD_glob_1m[[5]], 2)
  ),
  weight = rep(c(rep("unweighted", 3), "quadratic"), 3)
)
kable(kappa_1m, col.names = c("Accelerometer metric", "Intensity category", "Kappa", "Weight"), align = c("l", "l", "r", "l"))
```

### 6 months after surgery

```{r kappa_6m}
kappa_6m <- data.frame(
  acc_metric = c(rep("AC", 4), rep("MAD", 4), rep("ENMO", 4)),
  intens_cat = rep(c("Sedentary", "Moderate", "Vigorous", "Global"), 3),
  kappa = c(
    myround(kappa_AC_SED_6m[[5]], 2), myround(kappa_AC_MOD_6m[[5]], 2), myround(kappa_AC_VIG_6m[[5]], 2), myround(kappa_AC_glob_6m[[5]], 2),
    myround(kappa_ENMO_SED_6m[[5]], 2), myround(kappa_ENMO_MOD_6m[[5]], 2), myround(kappa_ENMO_VIG_6m[[5]], 2), myround(kappa_ENMO_glob_6m[[5]], 2),
    myround(kappa_MAD_SED_6m[[5]], 2), myround(kappa_MAD_MOD_6m[[5]], 2), myround(kappa_MAD_VIG_6m[[5]], 2), myround(kappa_MAD_glob_6m[[5]], 2)
  ),
  weight = rep(c(rep("unweighted", 3), "quadratic"), 3)
)
kable(kappa_6m, col.names = c("Accelerometer metric", "Intensity category", "Kappa", "Weight"), align = c("l", "l", "r", "l"))
```

### 12 months after surgery

```{r kappa_12m}
kappa_12m <- data.frame(
  acc_metric = c(rep("AC", 4), rep("MAD", 4), rep("ENMO", 4)),
  intens_cat = rep(c("Sedentary", "Moderate", "Vigorous", "Global"), 3),
  kappa = c(
    myround(kappa_AC_SED_12m[[5]], 2), myround(kappa_AC_MOD_12m[[5]], 2), myround(kappa_AC_VIG_12m[[5]], 2), myround(kappa_AC_glob_12m[[5]], 2),
    myround(kappa_ENMO_SED_12m[[5]], 2), myround(kappa_ENMO_MOD_12m[[5]], 2), myround(kappa_ENMO_VIG_12m[[5]], 2), myround(kappa_ENMO_glob_12m[[5]], 2),
    myround(kappa_MAD_SED_12m[[5]], 2), myround(kappa_MAD_MOD_12m[[5]], 2), myround(kappa_MAD_VIG_12m[[5]], 2), myround(kappa_MAD_glob_12m[[5]], 2)
  ),
  weight = rep(c(rep("unweighted", 3), "quadratic"), 3)
)
kable(kappa_12m, col.names = c("Accelerometer metric", "Intensity category", "Kappa", "Weight"), align = c("l", "l", "r", "l"))
```

## Percent agreement

```{r perc_agree_1m}
perc_agree <- data.frame(
  time_surgery = c(rep("1 month", 3), rep("6 months", 3), rep("12 months", 3)),
  acc_metric = rep(c("AC", "ENMO", "MAD"), 3),
  perc_agree = c(
    perc_agree_AC_ROC_1m, perc_agree_ENMO_ROC_1m, perc_agree_MAD_ROC_1m,
    perc_agree_AC_ROC_6m, perc_agree_ENMO_ROC_6m, perc_agree_MAD_ROC_6m,
    perc_agree_AC_ROC_12m, perc_agree_ENMO_ROC_12m, perc_agree_MAD_ROC_12m
  )
)
kable(perc_agree, col.names = c("Time after surgery", "Accelerometer metric", "Percent agreement"))
```
