---
title: "Cut-points analysis"
author: "Lucas Veras"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    number_sections: true
    toc_float: 
      collapsed: false
    theme: flatly
    code_folding: "show"
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load_packs}
library(here)
library(tidyverse)
```


# Sample

```{r load_data}
source(here("code", "scripts", "01_clean_data.R"))
```

```{r sample_message}
if (sample == 1) {
  message <- "participants who have been evaluated on at least 2 of the 3 evaluations (2nd, 3rd, 4th)"
} else {
  if (sample == 2) {
    message <- "participants who have been evaluated on the 3 eval (2nd, 3rd and 4th), with training sessions attendance <= 50%"
  } else {
    if (sample == 3) {
      message <- "participants who have been evaluated on the 3 eval (2nd, 3rd and 4th), no matter training sessions attendance"
    }
  }
}

n <- hip %>% 
  filter(!is.na(AC)) %>% 
  group_by(eval) %>% 
  summarise(
    n = n_distinct(ID)
  )
```

The selected sample is all `r message`. The number of participants for the 2nd, 3rd and 4th evaluations is `r n[1, 2]`, `r n[2, 2]` and `r n[3, 2]`, respectively.

Descriptive statistics for cardiorespiratory and accelerometry variables, separated by speed is show below, where `speed 0` is sitting rest, `speed 1` is standing rest, and the remaining `speed` values is the walking speed in km^.^h^-1^.

```{r explore}
source(here("code", "scripts", "02_explore.R"))
```

```{r cardio_acc_desc}
knitr::kable(
  cardio_acc_descriptives,
  col.names = c(
    "speed", "n",
    "kcal (mean)", "kcal (SD)",
    "VO[2] (mean)", "VO[2] (SD)",
    "MET (mean)", "MET (SD)",
    "AC (mean)", "AC (SD)",
    "ENMO (mean)", "ENMO (SD)",
    "MAD (mean)", "MAD (SD)"
  )
  )
```

# ROC curves

